Figure 6. Differential Protein expression

### 2022-01-06 reports/rmarkdown/manuscript/Figure6.Rmd
Figure 6. Differential Protein expression

---
title: "Differential Protein expression"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(rstatix)
  library(broom)
  library(niceRplots)
})  })  })

@@

```{r DPE Functions, message=FALSE, warning=FALSE, include=FALSE}
################################
# DEG's FUNCTION ACROSS GROUPS #
################################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
limma_dpe.fun <- function(group, confound) {
  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

@@

  # DPE across all groups
  top <- topTable(efit, adjust.method = "BH",number = "all",sort.by = "F") # ANOVA
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Antibody")

@@

@@

  if(layout == "multi") {
    layout(matrix(c(1,2,5,6,
                    7,7,8,8,
                    3,3,4,4),
           nrow = 3, ncol = 4,byrow = T),
           widths = c(.9,2,.9,2), heights = c(1,1,.01) ) }
  if(layout == "single") {
    layout(matrix(c(1,2),
                    nrow = 1, ncol = 2,byrow = T),widths = c(1,2) ) }

    ## Dendogram
    par(mgp=c(0,0.2,0), mar=c(4,.6,1,.2)) #(bottom, left, top, right) 
    plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .7 )
    #title(xlab="height", line=3)
    mtext( c("Height"), side=1, las = 1, cex = .7, line =1,xpd=T)

    if(gr == "Tissue_gr") {
      height <-1.5 
      add_letter("b")
    }else{height <- round(max(h$height))
      add_letter("a")}
    #lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob

    ## Heatmap
    par(mar=c(4,.2,1,2.5)) #(bottom, left, top, right) 
    image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
           col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
           breaks = breaks,
           axes=F)

    mtext( c("patient groups"), side=1, las = 1, cex = .7, line =1,xpd=T)

    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,0,0),
            y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39),
            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

@@

```{r Save stats to excel, include=FALSE}
# save TopTables pairwise
l_top_p <- list(Luminal = pairwise_dpe_L$top, Tissue = pairwise_dge_T$top)
imap(l_top_p, ~write.xlsx(.x,file=paste0("../../../results/DE/","dpe_pairwise_",.y,".xlsx"))) # Excel

# save TopTables across + cytokines
l_top_a <- list(Luminal = list(Across = dpe_L$top), Tissue = list(Across = dpe_T$top))
l <- list(l_top_a, stat, l_top_p)
names <- c("Luminal", "Tissue") %>% set_names()
p_list <- imap(names, ~flatten(map_depth(l, 1, .x)))

imap(p_list, ~write.xlsx(.x,file=paste0("../../../results/DE/","Proteins_",.y,".xlsx"))) # Excel

```{r Proteins plot, echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
figlabels <- letters

@@

par( mar=c(1.4,4,1.5,1)) #b,l,t,r 
violist(data=Luminal$sig_df,
          srt=45,
          genes= g,
          col = pal, transparency = 50,
          clustering=c(Luminal$groupings),smooth = 2 )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = .9)

add_letter("c")

@@

par( mar=c(24.5,5,1.5,1)) #b,l,t,r
violist(data=Tissue$sig_df,
          srt=45, 
          genes= rownames(Tissue$sig_df),
          col = pal, transparency = 50,
          clustering=Tissue$groupings, smooth = 2 )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = .9)

add_letter("d")

>>>>

---
title: "Figure 6. Differential Protein expression"
output: 
  pdf_document:
    fig_caption: yes
    keep_tex: yes
header-includes: 
- \usepackage{float}
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE}
##################
# LOAD LIBRARIES #
##################
#setwd("/Users/vilkal/work/Brolidens_work/Projects/Gabriella_repo/reports/rmarkdown/manuscript")
suppressWarnings({suppressMessages({suppressPackageStartupMessages({
  library(tidyverse)
  library(edgeR)
  library(openxlsx)
  library(rstatix)
  library(niceRplots)
})  })  })

@@

```{r DPE Functions, message=FALSE, warning=FALSE, include=FALSE}
################################
# DEG's FUNCTION ACROSS GROUPS #
################################
#group <- "Tissue_gr"
#group <- "Luminal_gr"
#group <- "L_T_groups"
limma_dpe.fun <- function(group, confound) {
  groups <- metadata[sample_use, group]
  n_gr <- length(na.omit(unique(groups)))

@@

  # DPE across all groups
  top <- topTable(efit, adjust.method = "BH",number = "all") # ANOVA
  top <- cbind(LogFC.intercept=0,top) # with intercept
  top <-  rownames_to_column(top, var = "Antibody")

@@

# L2 to T2/T3
metadata <- metadata %>% 
  mutate(L_T_groups = paste0(.$Luminal_gr, .$Tissue_gr)) %>%
  mutate(L_T_groups = ifelse(grepl("L2T2|L2T3", .$L_T_groups), .$L_T_groups, NA))

dpe_LT <- limma_dpe.fun("L_T_groups", confound=T)

@@

  if(layout == "multi") {
    layout(matrix(c(1,2,5,6,
                    7,7,8,8,
                    3,3,4,4),
           nrow = 3, ncol = 4,byrow = T),
           widths = c(1,2,.8,2), heights = c(1,1,.01) ) }
  if(layout == "single") {
    layout(matrix(c(1,2),
                    nrow = 1, ncol = 2,byrow = T),widths = c(1,2) ) }

    ## Dendogram
    par(mgp=c(0,0.2,0), mar=c(4,.6,2.5,.2)) #(bottom, left, top, right) 
    plot( rev(as.dendrogram(h)) , xlim=c(max(h$height),-.01),
          horiz = T, leaflab = "none",xaxs="i",yaxs="i",axes=T, cex.axis = .7 )
    #title(xlab="height", line=3)
    mtext( c("Height"), side=1, las = 1, cex = .7, line =1,xpd=T)

    if(gr == "Tissue_gr") {
      height <-1.5 
      add_letter("c")
    }else{height <- round(max(h$height))
      add_letter("a")}
    #lines(x=c(0,height), y=c(0.45,0.45), col="black", lty=1, lwd = 2.5) # only for grob

    ## Heatmap
    par(mar=c(4,.2,2.5,2.5)) #(bottom, left, top, right) 
    image( t(top_PROT_mfi[h$order,pull(df, "ID")][nrow(top_PROT_mfi):1,]),
           col = colorRampPalette(c("navy","navy","grey95","firebrick4","firebrick4") )(99),
           breaks = breaks,
           axes=F)
    mtext( "Differentialy Expressed Proteins", side=3, las = 1, cex = .7, line =1.5,xpd=T)
    mtext( c("Study Groups"), side=1, las = 1, cex = .7, line = 1,xpd=T)

    # Study group annotation
    end <- table(pal[factor(names(samples))[ order(df[1])]])
    end <- end[order(factor(names(end), levels = pal))]
    end <- map_dbl(cumsum(end), ~ (.x * (par("usr")[2])/length(samples)-.017) )
    start <- c(0, map_dbl(end[1:4], ~.x+0.015))

    axis(1, at = c(start, end), label = F, pos =1.03 , xpd=T, col="white",col.ticks="black")
    map2(start, end, ~lines(x=c(.y, .x), y= c(1.03, 1.03), xpd=T, cex=.8))
    # text labels
    t <- map2(start, end, ~(.x+.y)/2)
    map2(t,levels(factor(names(samples))), ~text(x=.x, y=1.04, .y,xpd=T, cex=.8))

    # Boxes 
    if(gr == "Luminal_gr") {
      lines(x=c(0,.603,.603,0,0),
            y=c(.488,.488,1.006, 1.006,.488), col="red", lty=1, lwd = 1.5)
      lines(x=c(.39,.94,.94,.39,.39),
            y=c(.488,.488,-0.008, -0.008,.488), col="red", lty=1, lwd = 1.5)}
    if(gr == "Tissue_gr") {
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,1.011, 1.011,.735), col="red", lty=1, lwd = 1.5)
      lines(x=c(.653,.95,.95,.653,.653),
            y=c(.735,.735,-0.012, -0.012,.735), col="red", lty=1, lwd = 1.5)
    }

@@

```{r Suppl. Tbl. 15 16 & 17, include=FALSE}
names <- c("15 Luminal", "16 Tissue") %>% set_names()

# save TopTables pairwise
l_top_p <- list(pairwise_dpe_L$top, pairwise_dge_T$top)  %>% set_names(names)

# save TopTables across + cytokines
l_top_a <- list( list(Across = dpe_L$top), list(Across = dpe_T$top))  %>% set_names(names)
l <- list(l_top_a, stat, l_top_p)
p_list <- imap(names, ~flatten(map_depth(l, 1, .x)))

imap(p_list, ~write.xlsx(.x,file=paste0("./Suppl.Tbl/","Suppl.Tbl.",.y,"_Proteins.xlsx"))) # Excel
write.xlsx(list("L2T2 vs L2T3"=dpe_LT$top),file=paste0("./Suppl.Tbl/","Suppl.Tbl.17"," L2T2_L2T3_Proteins.xlsx"))
```

```{r Figure 6., echo = FALSE, eval = TRUE, fig.height=10, fig.width=7}
figlabels <- letters

@@

par( mar=c(1.4,5,1.5,0)) #b,l,t,r 
violist(data=Luminal$sig_df,
          srt=45,
          genes= g,
          col = pal, transparency = 50,
          clustering=c(Luminal$groupings),smooth = 2 )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Luminal gr. Cytokines", line = 0.7, cex.main = .9)

rect(7, -0.5, 5.5, par("usr")[2], density = NULL, angle = 45, xpd = TRUE,
     col = "White", border = "White", lty = par("lty"), lwd = par("lwd")
     )

add_letter("b")

@@

par( mar=c(24.4,6,1.5,0)) #b,l,t,r
violist(data=Tissue$sig_df,
          srt=45, 
          genes= rownames(Tissue$sig_df),
          col = pal, transparency = 50,
          clustering=Tissue$groupings, smooth = 2 )
segments(x0 = stars$line_x0, x1 = stars$line_x1, y0 = stars$y_, y1 = stars$y_) 
text( stars$x , stars$y_+.05,
        labels = stars$p.adj.signif, srt = 0, xpd = TRUE, cex=1)
title(main = "Tissue gr. Cytokines", line = 0.7, cex.main = .9)

rect(6, -0.5, 4.5, par("usr")[2], density = NULL, angle = 45, xpd = TRUE,
     col = "White", border = "White", lty = par("lty"), lwd = par("lwd")
     )
add_letter("d")
